Title

### 2018-09-19 thesis/paper/paper.tex
Typos

\subsection{Representation of Wikidata as Facts}\label{subsec_representation}
To represent the content of Wikidata as facts the following predicates are introduced:
\begin{table}[H]
\caption{EDB predicates}\label{tab_predicates}
\begin{tabular}{llp{8.5cm}}
Predicate & Positions & Description \\
\hline
statement\_EDB & \makecell{statementID \\ entityID \\ propertyID \\ value (entityID or string)} & represents basic statements \\
\hline
qualifier\_EDB & \makecell{statementID \\ propertyID \\ value} & represents qualifiers on statements \\
\hline
reference\_EDB & \makecell{statementID \\ reference hash \\ propertyID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
\hline
item\_EDB & entityID & marks an entity as an item \\
property\_EDB & entityID & marks an entity as a property \\
\hline
unit\_EDB & \makecell{value \\ unitID} & denotes which unit a value has \\
\hline
rank\_EDB & \makecell{statementID \\ rank} & denotes the rank a statement has
\end{tabular}
\end{table}

@@

The task for each constraint is thus: Given the constraint and a constrained property, potentially with qualifiers, which rule finds all violations of the rule in a given set of triples, qualifiers and references if applicable? The given rule set should return all violations if queries with the following atoms:
\begin{align*}
&\text{violation\_statement}(S, E, P, V) \\
&\text{violation\_qualifier}(S, P, V) \\
&\text{violation\_reference}(S, H, P, V)
\end{align*}

@@

{\bf Rules:} Most constraints require more data than just the statements, qualifiers and references. This can include additional data from Wikidata to be provided as facts using the introduced predicates (see Table \ref{tab_predicates}), inequalities to be established as a pre-processing step (see Section \ref{sec_inequality}) or preparations to determine specific kinds of negation (see Section \ref{sec_statement_non-existence}). Afterwards, the rules necessary to determine violations are described.

@@

\subsection{Scope (Q53869507)}
\paragraph{Constraint}
The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier or reference.
\begin{table}[H]
\caption{property scope constraint (Q53869507)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statement, qualifier and reference \\
\hline
Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449) \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers or references} \\
\hline
Violation & constrained property in a position that is not allowed \\
\hline
\end{tabularx}
\end{table}

The required data specifies which data, in relation to the constrained statement, needs to be evaluated to conclude if the constraint was violated. The used abbreviations are listed in Table \ref{tab_abbreviations}. In this case this is simply each statement, qualifier or reference whose predicate is the constrained property.\\
The \emph{property scope} constraint requires the scope, meaning statement, qualifier or reference, to be set via qualifier \emph{property scope} (P5314). The cardinality is 1..3 because it is mandatory and has only three possible values.\\
It is violated if a statement, qualifier or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\

@@

\paragraph{Rules}
The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough. This means violations can be found using these rules:
\begin{align*}
\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{reference\_EDB}(S, H, \text{propertyConstant}, V)
\end{align*}
Which of these rules need to be used depends on the values of the \emph{property scope} qualifier. If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used. The case for \emph{as qualifier} and \emph{as reference} is analogue with the second and third rule, respectively.

@@

\subsection{Allowed entity types (Q52004125)}
\paragraph{Constraint}
The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property or a lexeme, although as of September 2018 lexemes are not deployed.

\begin{table}[H]
\caption{Allowed entity types constraint (Q52004125)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements and entity types\\
\hline
Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218) \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
\hline
Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The allowed entity types constraint is violated by a statement with the wrong entity type.  Constraints regarding the lexeme entity type can be ignored since they are currently not used. The Violations depend on a simple conjunction:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
\end{split}
\end{equation*}
The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}. The second rule works analogue for \emph{Wikibase property}.

@@

\begin{table}[H]
\caption{None of constraint (Q52558054)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements, qualifiers and references \\
\hline
Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier or reference} \\
\hline
Violation & \makecell{constrained statement, qualifier or reference with a forbidden value} \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain triples, qualifiers and references. The rules, depending on nothing but specific database entries, are
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.

\section{Constraints using Inequality}\label{sec_constraints_using_inequality}
All translatable constraints except the four covered in the previous Section require inequality in some way or another. This means it is necessary to extract additional information during pre-processing, namely the constants for which inequality needs to be stated. Each constraint will list the possible unequal entities in relation to the constrained statement. Multiple ways of implementing inequality while keeping the number of rules small will be discussed. Afterwards all constraints that can be translated using inequality will be explained.

@@

To save space and avoid redundancy the following rules will not explicitly state the inequality requirement predicate. Instead, assume that for every premise of the form \(B_1 \wedge \ldots\wedge B_m \wedge I_1 \wedge I_n\) where \(B_1 .. B_m\) are arbitrary atoms, \(I_1 .. I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), the rules 
\begin{equation*}
\text{req\_inequality}(t_{n_1}, t_{n_2})
 \leftarrow B_1 \wedge \ldots \wedge B_m
\end{equation*} for all \(n\) are introduced.\\

@@

\paragraph{Rules}
The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value. Since this would mark every constrained statement as a violations because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation*}
The rules for qualifiers and reference are analogue.

@@

By setting the \emph{constraint scope} (P4680) qualifier on the constraint it can be limited to only affect a subset of statement, qualifier, and reference. As of September 2018 this is not done on any constraint.

@@

\paragraph{Rules}
To handle this constraint inequality between all units need to be established. Then a violation can be identified by finding all statement\_EDB, qualifier\_EDB and reference\_EDB with the constrained property and a unit unequal to all allowed units.
The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement. To this end the units of a triple, qualifier or reference needs to be extracted during pre-processing and their inequality established. Quantities are rendered as string during pre-processing, specifying their value, bounds and unit.

To determine if a violation occurred, the unit of the constrained statement needs to be found. This is done via the unit-predicate as below
\begin{equation*}
\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit\_EDB}(V, U)
\end{equation*}
with qualifiers and references treated analogue.

@@

Violation are thus
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}

@@

\paragraph{Rules}
To represent a violation of the \emph{allowed qualifiers} constraint in datalog inequality between all specified allowed qualifiers and the properties of qualifiers is required. The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers. The notation for set inequality as introduced in Section \ref{subsec_3_allowed_units} will be used. With \(A\) the set of allowed qualifiers the rule is:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{qualifier\_EDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
\end{split}
\end{equation*}

\subsection{One-of (Q21510859)}
\paragraph{Constraint}
The \emph{one-of} constraint limits the values of a constraint statement, qualifier or reference to a set of given values.

\begin{table}[H]
\caption{One-of constraint (Q21510859)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements, qualifiers and references \\
\hline
Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
\hline
Violation & \makecell{constrained statement with a value unequal to all allowed values} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The one-of constraint fits in the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers and references. Inequality is established between all \emph{allowed values} and the values of statements, qualifiers and references, respectively. With \(A\) the set of allowed values the rules for triple, qualifier and reference are:
\begin{align*}
\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \\
\text{statement\_EDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{qualifier\_EDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{reference\_EDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{align*}

@@

If qualifier non-existence should be tested the predicates have to be adapted, since qualifiers do not have a unique identifier. To assign them a unique identity, statement ID, property and value have to be used. For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property and the value and next\_qualifier\([6]\) with ID, property and value for the current and next qualifier are introduced.\\

@@

For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge \ldots \wedge C_m\) with atoms \(C_i\) depending on the constraint which is used to test if the statement does not fulfil the requirement. Then rule one
\begin{equation*}
\text{require}(S, p, t) \leftarrow \text{first}(S, I) \wedge c
\end{equation*}
states that the statement requires the condition and
\begin{equation*}
\text{require}(S, p, t) \leftarrow \text{next}(O, S) \wedge \text{require}(O, p, t) \wedge c
\end{equation*}
propagates this down the statement chain.

@@

If the last statement of an item or the last qualifier of a statement has require the violation occurred. In the following constraints only the conjunction \(c\) and adjustments to this pattern will be described. The require-propagation is regarded as given.

@@

\paragraph{Rules}
The \emph{item requires statement} constraint requires inequality between all properties of statements and the \emph{required property} as well as between all statement values and the \emph{allowed values}. The order of the item statements needs to be established. An item violates the constraint if it a) has a statement using constrained property as predicate, b1) has no statement using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value. The required term \(t\) is the \emph{required property}. Without separators the conjunction is
\begin{equation*}
\begin{split}
c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
\end{split}
\end{equation*} \(\)
If a set of \emph{allowed values} \(A\) has been specified, a second set of require-rules have to be introduced with
\begin{equation*}
\begin{split}
c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{split}
\end{equation*}
Note that the first set of rules is still required, since the preceding statements do not need to have the \emph{required property}. In both cases violations can be found with
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \text{last}(O, I) \wedge{} \\
&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

\subsection{Value requires statement (Q21510864)}\label{subsec_value_requires_statement}
\paragraph{Constraint}
The \emph{value requires statement} constraint requires an additional statement on the value of constrained statement, possibly also using a specific value from the list of allowed values. Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.

@@

\subsection{Symmetric (Q21510862)}
The \emph{symmetric} constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value. It is thus equal to an \emph{inverse} constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and inverse property. It can be translated exactly as the inverse constraint if the inverse property where set to the same as the constrained property.

@@

\begin{example}
The \emph{cast member} (P161) property requires an item to have multiple statements using this property, since most plays or films have multiple cast members. This is an example for a constraint from which exception can and should be made and is only meant to point to possibly missing data.
\end{example}

>>>>

\subsection{Representation of Wikidata as Facts}\label{subsec_representation}
To represent the content of Wikidata as facts, the following predicates are introduced:
\begin{table}[H]
\caption{EDB predicates}\label{tab_predicates}
\begin{tabular}{llp{8.5cm}}
Predicate & Positions & Description \\
\hline
statement\_EDB & \makecell{statementID \\ entityID \\ propertyID \\ value (entityID or string)} & represents basic statements \\
\hline
qualifier\_EDB & \makecell{statementID \\ propertyID \\ value} & represents qualifiers on statements \\
\hline
reference\_EDB & \makecell{statementID \\ reference hash \\ propertyID \\ value} & represents references on statements; reference hash groups the property-value pairs belonging to one reference \\
\hline
item\_EDB & entityID & marks an entity as an item \\
property\_EDB & entityID & marks an entity as a property \\
\hline
unit\_EDB & \makecell{value \\ unitID} & denotes which unit a value has \\
\hline
rank\_EDB & \makecell{statementID \\ rank} & denotes the rank a statement has
\end{tabular}
\end{table}

@@

The task for each constraint is thus: Given the constraint and a constrained property, potentially with qualifiers, which rule finds all violations of the rule in a given set of triples, qualifiers and references if applicable? The given rule set should return all violations if queried with the following atoms:
\begin{align*}
&\text{violation\_statement}(S, E, P, V) \\
&\text{violation\_qualifier}(S, P, V) \\
&\text{violation\_reference}(S, H, P, V)
\end{align*}

@@

{\bf Rules:} Most constraints require more data than just the statements, qualifiers, and references. This can include additional data from Wikidata to be provided as facts using the introduced predicates (see Table \ref{tab_predicates}), inequalities to be established as a pre-processing step (see Section \ref{sec_constraints_using_inequality}), or preparations to determine specific kinds of negation (see Section \ref{sec_constraints_using_statement_non}). Afterwards, the rules necessary to determine violations are described.

@@

\subsection{Scope (Q53869507)}
\paragraph{Constraint}
The \emph{property scope} constraint specifies if a property may be used as predicate in a statement, qualifier, or reference.
\begin{table}[H]
\caption{property scope constraint (Q53869507)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statement, qualifier and reference \\
\hline
Qualifiers & \makecell{\emph{property scope} (P5314) -- 1..3 \\ must be \emph{as main value} (Q54828448), \emph{as qualifier} (Q54828449), \\ or \emph{as reference} (Q54828450) \\ specifies allowed use in statements, qualifiers, or references} \\
\hline
Violation & constrained property in a position that is not allowed \\
\hline
\end{tabularx}
\end{table}

The required data specifies which data, in relation to the constrained statement, needs to be evaluated to conclude if the constraint was violated. The used abbreviations are listed in Table \ref{tab_abbreviations}. In this case, this is simply each statement, qualifier, or reference whose predicate is the constrained property.\\
The \emph{property scope} constraint requires the scope, meaning statement, qualifier, or reference, to be set via qualifier \emph{property scope} (P5314). The cardinality is 1..3 because it is mandatory and has only three possible values.\\
It is violated if a statement, qualifier, or reference has a predicate with the \emph{property scope} constraint that does not allow this scope.\\

@@

\paragraph{Rules}
The property scope constraint requires no context information to be violated; the triple, qualifier, or reference with the constrained property is enough. This means violations can be found using these rules:
\begin{align*}
\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \text{statement\_EDB}(S, I, \text{propertyConstant}, V)\\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \text{qualifier\_EDB}(S, \text{propertyConstant}, V)\\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \text{reference\_EDB}(S, H, \text{propertyConstant}, V)
\end{align*}
Which of these rules need to be used depends on the values of the \emph{property scope} qualifier. If a concrete constrained on a property does not have \emph{as main value} as a value of the qualifier, the first rule is used. The case for \emph{as qualifier} and \emph{as reference} is analogous with the second and third rule, respectively.

@@

\subsection{Allowed entity types (Q52004125)}
\paragraph{Constraint}
The \emph{allowed entity types} constraint limits if a property may be used in a statement whose subject is an item, a property, or a lexeme, although as of September 2018 lexemes are not deployed.

\begin{table}[H]
\caption{Allowed entity types constraint (Q52004125)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements and entity types\\
\hline
Qualifiers & \makecell{allowed entity type (P2305) -- 1..3 \\ must be \emph{Wikibase item} (Q29934200), \emph{Wikibase property} (Q29934218), \\ or lexeme (Q51885771); \\ specifies the entity type the property may be used on} \\
\hline
Violation & \makecell{constrained statement on an entity of a type unequal to all allowed types} \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The allowed entity types constraint is violated by a statement with the wrong entity type.  Constraints regarding the lexeme entity type can be ignored since they are not currently used. The violations depend on a simple conjunction:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{item}(I)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{property}(I)
\end{split}
\end{equation*}
The first rule are created for all \emph{allowed entity type} constraints whose \emph{allowed entity type} does not contain \emph{Wikibase item}. The second rule works analogous for \emph{Wikibase property}.

@@

\begin{table}[H]
\caption{None of constraint (Q52558054)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements, qualifiers and references \\
\hline
Qualifiers & \makecell{\emph{forbidden value} (P2305) -- 1..* \\ values that may not be object of a constraint statement, qualifier, or reference} \\
\hline
Violation & \makecell{constrained statement, qualifier, or reference with a forbidden value} \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The none of constraint simply forbids certain values for properties, thus effectively prohibiting certain statements, qualifiers, and references. The rules, depending on nothing but specific database entries, are
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, \text{propertyConstant}, \text{forbiddenValueConstant}) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, \text{forbiddenValueConstant})
\end{split}
\end{equation*}
Given a specified constraint, each of the three rules is created once for each \emph{forbidden value}.

\section{Constraints using Inequality}\label{sec_constraints_using_inequality}
All translatable constraints except the four covered in the previous Section require inequality in some way or another. This means it is necessary to extract additional information during pre-processing, namely the constants for which inequality needs to be stated. Each constraint will list the possible unequal entities in relation to the constrained statement. Multiple ways of implementing inequality will be discussed. Afterwards all constraints that can be translated using inequality will be explained.

@@

To save space and avoid redundancy, the following rules will not explicitly state the inequality requirement predicate. Instead, assume that for every premise of the form \(B_1 \wedge \ldots\wedge B_m \wedge I_1 \wedge I_n\) where \(B_1 .. B_m\) are arbitrary atoms, \(I_1 .. I_n\) are atoms of the form \(\text{unequal}(t_{n_1}, t_{n_2})\) with \(t_{n_1}, t_{n_2}\) terms and all variables amount \(t_{n_1}\) and \(t_{n_2}\) appear in at least one of \(B_1 \ldots B_m\), the rules 
\begin{equation*}
\text{req\_inequality}(t_{n_1}, t_{n_2})
 \leftarrow B_1 \wedge \ldots \wedge B_m
\end{equation*} for all \(n\) are introduced.\\

@@

\paragraph{Rules}
The \emph{distinct values} constraint requires two distinct constrained statements with the same value. 
Two constrained statements with the same value can be found via a simple conjunction of two statement\_EDB atoms with the propertyConstant as predicate and a variable \(V\) as value. Since this would mark every constrained statement as a violation because it can fulfil both atoms at the same time, inequality of the statement IDs \(S\) and \(O\) must be tested as well.
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{statement\_EDB}(O, X, \text{propertyConstant}, V) \wedge{} \\
&\text{unequal}(S, O)
\end{split}
\end{equation*}
The rules for qualifiers and reference are analogous.

@@

By setting the \emph{constraint scope} (P4680) qualifier on the constraint, it can be limited to only affect a subset of statement, qualifier, and reference. As of September 2018 this is not done on any constraint.

@@

\paragraph{Rules}
To handle this constraint, inequality between all units need to be established. Then a violation can be identified by finding all statement\_EDB, qualifier\_EDB, and reference\_EDB with the constrained property and a unit unequal to all allowed units.
The allowed units constraint needs to determine inequality between all allowed units and each unit on a constrained statement. To this end the units of a statement, qualifier, or reference need to be extracted during pre-processing and their inequality established. Quantities are rendered as string during pre-processing, specifying their value, bounds, and unit.

To determine if a violation occurred, the unit of the constrained statement needs to be found. This is done via the unit-predicate as below
\begin{equation*}
\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge \text{unit\_EDB}(V, U)
\end{equation*}
with qualifiers and references treated analogous.

@@

Violations are thus
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_qualifier}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{qualifier\_EDB}(S, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
\text{violation\_reference}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{reference\_EDB}(S, H, \text{propertyConstant}, V) \wedge{} \\
&\text{unit}(V, U) \wedge \text{unequal}(\{A\}, U)
\end{split}
\end{equation*}

@@

\paragraph{Rules}
To represent a violation of the \emph{allowed qualifiers} constraint in datalog, inequality between all specified allowed qualifiers and the properties of qualifiers is required. The constraint is violated if a qualifier on a constrained statement exists that is unequal to all allowed qualifiers. The notation for set inequality as introduced in Section \ref{subsec_3_allowed_units} will be used. With \(A\) the set of allowed qualifiers the rule is:
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \\
&\text{qualifier\_EDB}(S, Q, O) \wedge \text{unequal}(\{A\}, Q)
\end{split}
\end{equation*}

\subsection{One-of (Q21510859)}
\paragraph{Constraint}
The \emph{one-of} constraint limits the values of a constraint statement, qualifier, or reference to a set of given values.

\begin{table}[H]
\caption{One-of constraint (Q21510859)}
\begin{tabularx}{\textwidth}{ ll X}
\hline
Required data & constrained statements, qualifiers, and references \\
\hline
Qualifiers & \makecell{\emph{allowed values} (P2305) -- 1..* \\ values that may be used as objects of a constrained statement} \\
\hline
Violation & \makecell{constrained statement with a value unequal to all allowed values} \\
\hline
Rule features & inequality \\
\hline
\end{tabularx}
\end{table}

@@

\paragraph{Rules}
The one-of constraint fits the pattern established by the previous two constraints and disallows certain property-value combinations in statements, qualifiers, and references. Inequality is established between all \emph{allowed values} and the values of statements, qualifiers, and references, respectively. With \(A\) the set of allowed values the rules for triple, qualifier, and reference are:
\begin{align*}
\text{violation\_statement}(S, I, \text{propertyConstant}, V) &\leftarrow \\
\text{statement\_EDB}(S, I, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_qualifier}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{qualifier\_EDB}(S, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V) \\
\text{violation\_reference}(S, \text{propertyConstant}, V) &\leftarrow \\
\text{reference\_EDB}(S, H, &\text{propertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{align*}

@@

If qualifier non-existence should be tested, the predicates have to be adapted since qualifiers do not have a unique identifier. To assign them a unique identity, statement ID, property, and value have to be used. For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property, and the value, as well as next\_qualifier\([6]\) with ID, property, and value for the current and next qualifier are introduced.\\

@@

For statements the pattern consists of two rules, both utilising the same conjunction \(c = C_1 \wedge \ldots \wedge C_m\) with atoms \(C_i\) depending on the constraint. This conjunction is used to test if the statement does not fulfil the requirement. Then rule one
\begin{equation*}
\text{require}(S, p, t) \leftarrow \text{first}(S, I) \wedge c
\end{equation*}
states that the statement requires the condition and
\begin{equation*}
\text{require}(S, p, t) \leftarrow \text{next}(O, S) \wedge \text{require}(O, p, t) \wedge c
\end{equation*}
propagates this down the statement chain.

@@

If the last statement of an item or the last qualifier of a statement has the require-predicate the violation occurred. In the following constraints only the conjunction \(c\) and adjustments to this pattern will be described. The require-propagation is regarded as given.

@@

\paragraph{Rules}
The \emph{item requires statement} constraint requires inequality between all properties of statements and the \emph{required property} as well as between all statement values and the \emph{allowed values}. The order of the item statements needs to be established. An item violates the constraint if it a) has a statement using the constrained property as predicate, b1) has no statements using the \emph{required property} as predicate, or (if \emph{allowed values} are specified) b2) has only statement using the \emph{required property} as predicate that do not use one of the \emph{allowed values} as value. The required term \(t\) is the \emph{required property}. Without separators the conjunction is
\begin{equation*}
\begin{split}
c_1 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, P, V) \wedge \text{unequal}(\text{requiredPropertyConstant}, P)
\end{split}
\end{equation*} \(\)
If a set of \emph{allowed values} \(A\) has been specified, a second set of require-rules have to be introduced with
\begin{equation*}
\begin{split}
c_2 = &\text{statement\_EDB}(Q, I, \text{propertyConstant}, X) \wedge{} \\
&\text{statement\_EDB}(S, I, \text{requiredPropertyConstant}, V) \wedge \text{unequal}(\{A\}, V)
\end{split}
\end{equation*}
Note that the first set of rules is still required, since the preceding statements do not need to have the \emph{required property}. In both cases violations can be found with
\begin{equation*}
\begin{split}
\text{violation\_statement}&(S, I, \text{propertyConstant}, V) \leftarrow \\
&\text{statement\_EDB}(S, I, \text{propertyConstant}, V) \wedge{} \text{last}(O, I) \wedge{} \\
&\text{require}(O, \text{propertyConstant}, \text{requiredPropertyConstant})
\end{split}
\end{equation*}

\subsection{Value requires statement (Q21510864)}\label{subsec_value_requires_statement}
\paragraph{Constraint}
The \emph{value requires statement} constraint requires an additional statement on the value of the constrained statement, possibly also using a specific value from the list of allowed values. Since the constraint works very similar to item requires statement (see \ref{subsec_item_requires_statement}) --- requiring a statement on the value of the constrained statement instead of on the item --- Table \ref{tab_item_requires_statements} is not repeated.

@@

\subsection{Symmetric (Q21510862)}
The \emph{symmetric} constraint requires an additional statement on the value of a constrained statement with the constrained property and the constrained item as value. It is thus equal to an \emph{inverse} constraint (see \ref{subsec_2_inverse}) with the same property as constrained property and inverse property. It can be translated exactly as the inverse constraint if the inverse property were set to the same as the constrained property.

@@

\begin{example}
The \emph{cast member} (P161) property requires an item to have multiple statements using this property, since most plays or films have multiple cast members. This is an example for a constraint from which exceptions can and should be made and is only meant to point to possibly missing data.
\end{example}
